---
title: "HW3"
output: html_document
date: "2023-10-04"
---
## Question 1
### Paths
If A wins the game, there are 10 different paths for A to win.
I assume that the outcome of one path doesn't influence or depend on any other path.

#### Path type 1- A wins all the first three sets and game ends 
1. A wins set 1, set 2, set 3

#### Path type 2- A wins three sets out of the first four sets and game ends 
1. A wins set 1, set 2, set 4
2. A wins set 1, set 3, set 4
3. A wins set 2, set 3, set 4

#### Path type 3- A wins three sets out of the five sets 
1. A wins set 1, set 2, set 5
2. A wins set 1, set 3, set 5
3. A wins set 1, set 4, set 5
4. A wins set 2, set 3, set 5
5. A wins set 2, set 4, set 5
6. A wins set 3, set 4, set 5

### Probability of following each path to win 
From \(P_A=0.6\) (which is the probability that team A wins a set), we know \(P_B=1-0.6=0.4\) (which is the probability that team B wins a set). I assume that both of the probabilities are constant and following any specific path is independent of the other paths.

#### For following any path from path type 1 to win
A wins the first 3 sets. The probability is \(0.6^3*0.4^0\)

#### For following any path from path type 2 to win
A wins 3 sets out of the first 4 sets, which means B wins 1 set.The probability is \(0.6^3*0.4^1\)

#### For following any path from path type 3 to win
A wins 3 sets out of the 5 sets, which means B wins 2 sets. The probability is \(0.6^3 *0.4^2 \)

### Overall Probability 
The total probability is \(0.6^3*0.4^0*1+0.6^3*0.4^1*3+0.6^3 *0.4^2*6=0.68256\)

```{r}
y <- 0.6^3 * 0.4^0 * 1 + 
     0.6^3 * 0.4^1 * 3 + 
     0.6^3 * 0.4^2 * 6
print(y)
```

## Question 2
Before that, we need to know the probability that B wins the first game. Similar to question 1:
```{r}
y <- 0.4^3 * 0.6^0 * 1 + 
     0.4^3 * 0.6^1 * 3 + 
     0.4^3 * 0.6^2 * 6
print(y)
```
### Paths
If A wins the championship, there are 10 different paths for A to win. 
I assume that the outcome of one path doesn't influence or depend on any other path.

#### Path type 1- A wins all the first three games
1. A wins game 1, game 2, game 3

#### Path type 2- A wins three out of the first four games
1. A wins game 1, game 2, game 4
2. A wins game 1, game 3, game 4
3. A wins game 2, game 3, game 4

#### Path type 3- A wins three out of the five games 
1. A wins game 1, game 2, game 5
2. A wins game 1, game 3, game 5
3. A wins game 1, game 4, game 5
4. A wins game 2, game 3, game 5
5. A wins game 2, game 4, game 5
6. A wins game 3, game 4, game 5

We know the probability that A wins a game is \( 0.68256\) and the probability that B wins a game is \( 0.31744\) 

#### For following any path from path type 1 to win
A wins the first 3 games. The probability is \(0.68256^3\)

#### For following any path from path type 2 to win
A wins 3 sets out of the first 4 games, which means B wins 1 game. The probability is \(0.68256^3*0.31744^1\)

#### For following any path from path type 3 to win
A wins 3 sets out of the 5 games, which means B wins 2 games. The probability is \(0.68256^3 *0.31744^2 \)

### Overall Probability
The overall probability is: 
```{r}
y <- 0.68256^3 * 1 + 
     0.68256^3 * 0.31744^1 * 3 + 
     0.68256^3 * 0.31744^2 * 6
print(y)
```

## Question 3
### Calculate the probability that A wins a game
If A wins the game, there are 10 different paths for A to win. 

#### Path type 1- A wins all the first three sets and game ends 
1. A wins set 1, set 2, set 3

#### Path type 2- A wins three sets out of the first four sets and game ends 
1. A wins set 1, set 2, set 4
2. A wins set 1, set 3, set 4
3. A wins set 2, set 3, set 4

#### Path type 3- A wins three sets out of the five sets 
1. A wins set 1, set 2, set 5
2. A wins set 1, set 3, set 5
3. A wins set 1, set 4, set 5
4. A wins set 2, set 3, set 5
5. A wins set 2, set 4, set 5
6. A wins set 3, set 4, set 5

#### For following any path from path type 1 to win
A wins the first 3 sets.The probability is \(X^3*(1-X)^0\)

#### For following any path from path type 2 to win
A wins 3 sets out of the first 4 sets, which means B wins 1 set.The probability is \(X^3*(1-X)^1\)

#### For following any path from path type 3 to win
A wins 3 sets out of the 5 sets, which means B wins 2 sets. The probability is \(X^3 *(1-X)^2 \)

#### Overall Probability
The total probability a is:
```{r}
a <- function(x){
  return(x^3 * (1-x)^0 * 1 + x^3 * (1-x)^1 * 3+ x^3 * (1-x)^2 * 6)
}
```

### Calculate the probability that B wins a game
If B wins the game, there are 10 different paths for B to win. 

#### Path type 1- B wins all the first three sets and game ends 
1. B wins set 1, set 2, set 3

#### Path type 2- B wins three sets out of the first four sets and game ends 
1. B wins set 1, set 2, set 4
2. B wins set 1, set 3, set 4
3. B wins set 2, set 3, set 4

#### Path type 3- B wins three sets out of the five sets 
1. B wins set 1, set 2, set 5
2. B wins set 1, set 3, set 5
3. B wins set 1, set 4, set 5
4. B wins set 2, set 3, set 5
5. B wins set 2, set 4, set 5
6. B wins set 3, set 4, set 5

#### For following any path from path type 1 to win
B wins the first 3 sets.The probability is \((1-X)^3*X^0\)

#### For following any path from path type 2 to win
B wins 3 sets out of the first 4 sets, which means A wins 1 set.The probability is \((1-X)^3*X^1\)

#### For following any path from path type 3 to win
B wins 3 sets out of the 5 sets, which means A wins 2 sets. The probability is \((1-X)^3 *X^2 \)

#### Overall Probability
The total probability b is:
```{r}
b <- function(x){
  return((1-x)^3 * x^0 * 1+ (1-x)^3 * x^1 * 3+ (1-x)^3 * x^2 * 6)
}
```

#### Calculate the probability that A wins the championship 
If A wins the championship, there are 10 different paths for A to win. 

#### Path type 1- A wins all the first three games
1. A wins game 1, game 2, game 3

#### Path type 2- A wins three out of the first four games
1. A wins game 1, game 2, game 4
2. A wins game 1, game 3, game 4
3. A wins game 2, game 3, game 4

#### Path type 3- A wins three out of the five games 
1. A wins game 1, game 2, game 5
2. A wins game 1, game 3, game 5
3. A wins game 1, game 4, game 5
4. A wins game 2, game 3, game 5
5. A wins game 2, game 4, game 5
6. A wins game 3, game 4, game 5

The probability that A wins the game is a. The probability that B wins the game is b. 

#### For following any path from path type 1 to win
A wins the first 3 games.The probability is \(a^3\)

#### For following any path from path type 2 to win
A wins 3 games out of the first 4 games, which means B wins 1 game. The probability is \(a^3*b^1\)

#### For following any path from path type 3 to win
A wins 3 games out of the 5 games, which means B wins 2 games. The probability is \(a^3 *b^2 \)

### Overall Probability
The overall probability is: 

```{r}
a <- function(x){
  return(x^3 * (1-x)^0 * 1 + x^3 * (1-x)^1 * 3+ x^3 * (1-x)^2 * 6)
}

b <- function(x){
  return((1-x)^3 * x^0 * 1+ (1-x)^3 * x^1 * 3+ (1-x)^3 * x^2 * 6)
}

y <- function(x){
  return(a(x)^3 * b(x)^0 * 1 + a(x)^3 * b(x)^1 * 3 + a(x)^3 * b(x)^2 * 6)
}

x_values <- seq(0,1, by=0.01)
y_values <- sapply(x_values,y)

plot(x_values,y_values,main="Probability that A Wins the Championship",xlab="Probability that A Wins in any Given Set",ylab="Probability that A Wins the Championship")
```

## Question 4
### Interpreting what we need to do using Bayes Law 
\[P(P_B=0.55|\text{Team A wins the championship}) = \frac{P(\text{Team A wins the championship})|P_B=0.55) \times P(P_B=0.55)}{P(\text{Team A wins the championship})}\]

### Probability that team A wins the championship
The assumption is either \( P_B=0.55\) or \( P_B=0.65\); therefore, there are two scenarios. The possibility of each one is 0.5.

#### Scenario 1
If \( P_B=0.55\), then \( P_A=0.45\)
The possibility that A wins the championship is \(0.3293834\)

```{r}
a <- function(x){
  return(x^3 * (1-x)^0 * 1 + x^3 * (1-x)^1 * 3+ x^3 * (1-x)^2 * 6)
}

b <- function(x){
  return((1-x)^3 * x^0 * 1+ (1-x)^3 * x^1 * 3+ (1-x)^3 * x^2 * 6)
}

y <- function(x){
  return(a(x)^3 * b(x)^0 * 1 + a(x)^3 * b(x)^1 * 3 + a(x)^3 * b(x)^2 * 6)
}

y(0.45)
```

#### Scenario 2
If \( P_B=0.65\), then \( P_A=0.35\)
The possibility that A wins the championship is \( 0.08849626\)

```{r}
a <- function(x){
  return(x^3 * (1-x)^0 * 1 + x^3 * (1-x)^1 * 3+ x^3 * (1-x)^2 * 6)
}

b <- function(x){
  return((1-x)^3 * x^0 * 1+ (1-x)^3 * x^1 * 3+ (1-x)^3 * x^2 * 6)
}

y <- function(x){
  return(a(x)^3 * b(x)^0 * 1 + a(x)^3 * b(x)^1 * 3 + a(x)^3 * b(x)^2 * 6)
}

y(0.35)
```
#### Overall Probability
The possibility for both of the two scenarios is 0.5; therefore, the overall probability that team A wins the championship is
```{r}
0.5 * 0.3293834 + 0.5 * 0.08849626
```

### Interpreting what we need to do using Bayes Law- Continue 
\[P(P_B=0.55|\text{Team A wins the championship}) = \frac{P(\text{Team A wins the championship})|P_B=0.55) \times P(P_B=0.55)}{P(\text{Team A wins the championship})}=\frac{0.3293834 \times0.5}{0.2089398}=0.7882256\]
```{r}
0.3293834 * 0.5/ 0.2089398
```

## Question 5
If \( P_A=0.55\), then \( P_A=0.45\).
Let's say the length is \( 2k+1,k∈Z+\)

### If we assume each game has (2k+1) sets:  
### When k = 1, length is 3:
The probability that Team A wins a game is \({2 \choose 2} * 0.55^2 * 0.45^0 + {2 \choose 1} * 0.55^2 * 0.45^1\)  
1. \({2 \choose 2} * 0.55^2 * 0.45^0 \) is the total probability for following any paths in path type 1: A wins all the first two sets and game ends; \( {2 \choose 2}\) is the number of paths for this type.     
2. \({2 \choose 1} * 0.55^2 * 0.45^1 \) is the total probability for following any paths in path type 2: A and B wins half of the first two sets and A wins the last set; \( {2 \choose 1}\) is the number of paths for this type.   

### When k = 2, length is 5:
The probability that Team A wins a game is \({3 \choose 3} * 0.55^3 * 0.45^0 + {3 \choose 2} * 0.55^3 * 0.45^1 + {4 \choose 2} * 0.55^3 * 0.45^2\)    
1. \({3 \choose 3} * 0.55^3 * 0.45^0 \) is the total probability for following any paths in path type 1: A wins all the first three sets and game ends; \( {3 \choose 3}\) is the number of paths for this type.   
2. \({3 \choose 2} * 0.55^3 * 0.45^1 \) is the total probability for following any paths in path type 2: A wins three out of the first four sets; \( {3 \choose 2}\) is the number of paths for this type.   
3. \({4 \choose 2} * 0.55^3 * 0.45^2 \) is the total probability for following any paths in path type 3: A and B wins half of the first four sets and A wins the last set; \( {4 \choose 2}\) is the number of paths for this type.   

### When k = 3, length is 7:
The probability that Team A wins a game is \({4 \choose 4} * 0.55^4 * 0.45^0 + {4 \choose 3} * 0.55^4 * 0.45^1 + {5 \choose 3} * 0.55^4 * 0.45^2 +  {6 \choose 3} * 0.55^4 * 0.45^3\)  
1. \({4 \choose 4} * 0.55^4 * 0.45^0 \) is the total probability for following any paths in path type 1: A wins all the first four sets and game ends; \( {4 \choose 4}\) is the number of paths for this type.   
2. \({4 \choose 3} * 0.55^4 * 0.45^1 \) is the total probability for following any paths in path type 2: A wins four out of the first five sets; \( {4 \choose 3}\) is the number of paths for this type.   
3. \({5 \choose 3} * 0.55^4 * 0.45^2 \) is the total probability for following any paths in path type 3: A wins four out of the first six sets; \( {5 \choose 3}\) is the number of paths for this type.   
4. \({6 \choose 3} * 0.55^4 * 0.45^3 \) is the total probability for following any paths in path type 4: A and B wins half of the first six sets and A wins the last set; \( {6 \choose 3}\) is the number of paths for this type.     

### To sum, when length is 2k + 1 
The probability that Team A wins a game is \({k+1 \choose k+1} * 0.55^{k+1}* 0.45^0 + {k+1 \choose k} * 0.55^{k+1} * 0.45^1 + {k+2 \choose k} * 0.55^{k+1} * 0.45^2 + ... + {2k \choose k} * 0.55^{k+1} * 0.45^k \)

```{r}
c <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.55^(k+1) * 0.45^i
  }
  
  return(sum_term)
}
c(2) #test 

```
We only need to replace 0.55 with 0.45 to get the probability that Team B wins a game: 
```{r}
d <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.45^(k+1) * 0.55^i
  }
  
  return(sum_term)
}
d(2) #test 
```

We only need to replace 0.55 with c(k) (the probability that Team A wins a game) to get the probability that Team A wins a champion: 
```{r}
c <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.55^(k+1) * 0.45^i
  }
  
  return(sum_term)
}

d <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.45^(k+1) * 0.55^i
  }
  
  return(sum_term)
}

e <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * c(k)^(k+1) * d(k)^i
  }
  
  return(sum_term)
}
e(5) #test 
```

Therefore, if we want e(k)>=0.85
```{r}
c <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.55^(k+1) * 0.45^i
  }
  
  return(sum_term)
}

d <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * 0.45^(k+1) * 0.55^i
  }
  
  return(sum_term)
}

e <- function(k) {
  sum_term <- 0
  
  for (i in 0:k) {
    sum_term <- sum_term + choose(k + i, k) * c(k)^(k+1) * d(k)^i
  }
  
  return(sum_term)
}

k <- 1
while (e(k) < 0.85) {
  k <- k + 1
}

print(k)
```
The shortest final series length will be \(2*6+1 = 13\)

### If we assume each game has 5 sets: 
The possibility that A wins the game is \( 0.5931269\)
```{r}
a <- function(x){
  return(x^3 * (1-x)^0 * 1 + x^3 * (1-x)^1 * 3+ x^3 * (1-x)^2 * 6)
}

a(0.55)
```
The possibility that B wins the game is \( 0.4068731\)
```{r}
b <- function(x){
  return((1-x)^3 * x^0 * 1+ (1-x)^3 * x^1 * 3+ (1-x)^3 * x^2 * 6)
}

b(0.55)
```
Therefore, if we want g(k)>=0.85
```{r}
g <- function(k) {
  sum <- 0
  for (i in 0:k) {
    sum <- sum + choose(k + i, k) * 0.5931269^(k+1) * 0.4068731^i
  }
  return(sum)
}

k <- 1
while (g(k) < 0.85) {
  k <- k + 1
}

print(k)
```
The shortest final series length will be \(2*15+1 = 31\)